---
title: "Chaudhary_homework_8"
author: "Jyoti Chaudhary"
date: "November 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(tidyr)
library(stringr)
```


1. Write a set of functions to clean a column in the WHO data from last week. For example, use a modification (no need for commas in the output) of the clean_who_number_spaces() function from class, other functions to pull out the upper and lower bounds, and a function  that calls all of those functions in order to take as input a column in the WHO format and output a 3 column data frame. Name that last function "fix_who_column". 


```{r message=FALSE, warning=FALSE}

clean_who_number_spaces <- function(x){
  ## Change formating of numbers from xxx xxx to xxxxxx
  
  clean_triplet_pair <- 
    function(x) str_replace(x, "([0-9]{1,3}) ([0-9]{3})", "\\1\\2")
  number_of_iterations <- max(map_int(x, ~ str_count(., " ")))
  for (i in 1:number_of_iterations){
    x <- clean_triplet_pair(x)
  }
}

fix_who_column <- function(df, col1, col2, col3){
  
  ## Take col1 as input and output 3 column dataframe
  
  library(tidyverse)
  library(stringr)
  library(dplyr)
  
  pos_openbraces <- str_locate(df[[col1]], "\\[")
  pos_closebraces <- str_locate(df[[col1]], "\\]")
  pos_dash <- str_locate(df[[col1]], "-")
  
  pos_openbraces <- pos_openbraces[,1]
  pos_closebraces <- pos_closebraces[,1]
  pos_dash <- pos_dash[,1]
  
  
  df[[col2]] <- ifelse(pos_openbraces != "NA", str_sub(df[[col1]], pos_openbraces + 1, pos_dash - 1), "NA")
  
  df[[col3]] <- ifelse(pos_openbraces != "NA", str_sub(df[[col1]], pos_dash + 1, pos_closebraces - 1), "NA")
  
  df[[col1]] <- ifelse(pos_openbraces != "NA", str_sub(df[[col1]], end = pos_openbraces - 1), "NA")
  
  clean_who_number_spaces(df[[col1]])   
  clean_who_number_spaces(df[[col2]])
  clean_who_number_spaces(df[[col3]])
  
  result <- data.frame("Patient.count" = df[[col1]], "lower.bound" = df[[col2]], "upper.bound" = df[[col3]])
  result
}

```

Read "data.csv" file and apply fix_who_column

```{r }

## Read "Number of People living with HIV" dataset
HIV_data <- read_csv("data.csv", skip=1)

colnames(HIV_data) <- c("Region", "Patient.count")

## Call function to read "Patient.count" as input and output a dataframe with 3 columns

fix_who_column(HIV_data, "Patient.count", "lower.bound", "upper.bound")

```