---
title: "Chaudhary_homework_7"
author: "Jyoti Chaudhary"
date: "November 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(tidyr)
library(rvest)
library(readxl)
library(stringr)
library(broom)
library(ggplot2)
library(RColorBrewer)
library(viridis)
library(ggthemes)
library(choroplethr)
library(choroplethrMaps)

```

## TASK 1
Download the csv file on maternal mortality at   http://apps.who.int/gho/athena/data/xmart.csv?target=GHO/MDG_0000000025,MDG_0000000026&profile=crosstable&filter=COUNTRY:*;YEAR:*&x-sideaxis=COUNTRY;YEAR&x-topaxis=GHO and store it in a data directory in your project repo. (You can read about the data  at http://apps.who.int/gho/data/view.main.SDG31v?lang=en ) .Although the data is tidy, it needs cleaning. Notice that spaces are used instead of commas in numbers. Turn the numbers for maternal mortality inside square brackets into two new variables, lower_bound and upper_bound

```{r message=FALSE, warning=FALSE}
xmart <- read_csv("xmart.csv")

colnames(xmart) <- c("Country", "Year", "Mortality.ratio", "Birth.attended")

xmart$Year <- str_sub(xmart$Year, end = 4)

xmart$pos_openbraces <- str_locate(xmart$Mortality.ratio, "\\[")
xmart$pos_closebraces <- str_locate(xmart$Mortality.ratio, "\\]")
xmart$pos_dash <- str_locate(xmart$Mortality.ratio, "-")


xmart$pos_openbraces <- xmart$pos_openbraces[,1]
xmart$pos_closebraces <- xmart$pos_closebraces[,1]
xmart$pos_dash <- xmart$pos_dash[,1]


xmart$lower_bound <- ifelse(xmart$pos_openbraces != "NA", str_sub(xmart$Mortality.ratio, xmart$pos_openbraces + 1, xmart$pos_dash - 1), "NA")

xmart$upper_bound <- ifelse(xmart$pos_openbraces != "NA", str_sub(xmart$Mortality.ratio, xmart$pos_dash + 1, xmart$pos_closebraces -1), "NA")

xmart$Mortality.ratio <- ifelse(xmart$pos_openbraces != "NA", str_sub(xmart$Mortality.ratio, end = xmart$pos_openbraces - 1), "NA")

xmart$Mortality.ratio <- as.integer(str_replace_all(xmart$Mortality.ratio, " ", ""))
xmart$lower_bound <- as.integer(str_replace_all(xmart$lower_bound, " ", ""))
xmart$upper_bound <- as.integer(str_replace_all(xmart$upper_bound, " ", ""))

xmart_clean <- select(xmart, Country, Year, Mortality.ratio, Birth.attended, lower_bound, upper_bound)

head(xmart_clean)

```

## TASK 2
Download the gross national income data from http://apps.who.int/gho/data/node.main.GNI107?lang=en (Download the complete data as a "CVS table", storing it in the data folder).
Tidy this data, so that there is one entry per country per year

```{r message=FALSE, warning=FALSE}

income_data <- read_csv("data.csv", col_names = TRUE, col_types = NULL, skip = 1)
tall_income_data <- gather(income_data, key="Year", value = "Income", 2:25)
tall_income_data$Income <-  as.integer(str_replace_all(tall_income_data$Income, " ", ""))
clean_income_data <- filter(tall_income_data, tall_income_data$Income != "NA")

# Remove duplicate rows on Country and Year
clean_income_data <- clean_income_data[!duplicated(clean_income_data[,c('Country', 'Year')]),]

colnames(clean_income_data) <- c("COUNTRY", "YEAR", "INCOME")

head(clean_income_data)

```

## TASK 3
Merge the two files by country and year, retaining only rows for which you have per capita income and either maternal mortality or percentage attended births

```{r message=FALSE, warning=FALSE}

merge_data <- merge(clean_income_data, xmart_clean, by.x = c("COUNTRY", "YEAR"), by.y = c("Country", "Year")) %>% filter(INCOME != "NA" & (Mortality.ratio != "NA" | Birth.attended != "NA"))

head(merge_data)

```

## TASK 4
Make a two polished  and informative graphs, one relating income to maternal mortality and the other relating income to percentage attended births
Save a csv file with the merged, cleaned, and tidy data, using a suitable name of your choosing in the data folder.

```{r message=FALSE, warning=FALSE}

mortality_data <- merge_data %>% filter(Mortality.ratio != "NA")

ggplot(data = mortality_data, 
       aes(x = INCOME, y = Mortality.ratio)) +
  geom_point(alpha = 0.6, size = 3) +
  xlab("Gross national income per capita (dollars)") +
  ylab("Maternal mortality ratio (per 100 000 live births)") + ylim(c(0, 1500)) + xlim(c(0, 50000))


birthattended_data <- merge_data %>% filter(Birth.attended != "NA")


ggplot(data = birthattended_data, 
       aes(x = INCOME, y = Birth.attended)) +
  geom_point(alpha = 0.6, size = 3) +
  xlab("Gross national income per capita (dollars)") +
  ylab("Births attended by skilled health personnel (%)") + xlim(c(0, 75000))

```

## TASK 5
Use the country_choropleth() function in the choroplethr package to make a world map of maternal mortality, using the most recent year for each country in the merged dataset you created. The defaults of the function will be fine; no need to tune up the plot. You can read the help file to see how the data must be formatted; you may need to rename the countries from the dataset that you've created.


```{r message=FALSE, warning=FALSE}

country_year <-mortality_data %>%
        group_by(COUNTRY) %>% arrange(desc(YEAR)) %>%
        ddply("COUNTRY", summarise, YEAR = max(YEAR)) 

country_mortality <- left_join(country_year, mortality_data, by=c("COUNTRY" = "COUNTRY", "YEAR" = "YEAR")) %>%  select(COUNTRY, Mortality.ratio)

country_mortality$COUNTRY <- str_to_lower(country_mortality$COUNTRY)
        
colnames(country_mortality) <- c("region", "value")

country_choropleth(country_mortality, 
                 title = "World Map Of Maternal Mortality", 
                 legend = "Maternal Mortality Ratio", num_colors = 9)

```
