---
title: "Chaudhary_week_5b"
author: "Jyoti Chaudhary"
date: "October 30, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(plyr)
library(stringr)
library(tidyr)
library(dplyr)
library(broom)

```

## PROBLEM 1
3) Consider the following three websites,
    Medal count from Rio Olympics
        http://www.cbssports.com/olympics/news/2016-rio-olympics-medal-tracker/
    Three letter contry codes:
               https://en.wikipedia.org/wiki/List_of_IOC_country_codes
        (Note: I originally posted the page http://www.nationsonline.org/oneworld/country_code_list.htm , but that has ISO codes, not IOC codes)
    Populations of countries:
        https://en.wikipedia.org/wiki/List_of_countries_by_population_(United_Nations)
a) Please create a dataset with variables equal to the country name, the UN continental region, the population (2016), and the number of gold, silver, bronze, and total medals from the Rio Olympics. 


```{r message=FALSE, warning=FALSE}

# Extracting olympic medal table from the webpage

url1 <- "http://www.cbssports.com/olympics/news/2016-rio-olympics-medal-tracker/"

olympic_medal <- url1 %>% read_html() %>% html_nodes("table") %>%
  html_table(fill = TRUE) 

olympic_medal <- data.frame(olympic_medal)

colnames(olympic_medal) <- c("COUNTRY", "GOLD", "SILVER", "BRONZE", "TOTAL")

# Extracting Country Codes table from webpage. 3 Tables extracted and combined in one 

url2 <- "https://en.wikipedia.org/wiki/List_of_IOC_country_codes"

Country_code <- url2 %>% read_html() %>% html_nodes("table") %>% .[1:3] %>%
  html_table(fill = TRUE) 


code_df1 <- data.frame(Country_code[[1]]$Code, Country_code[[1]]$`Nation (NOC)`)
code_df2 <- data.frame(Country_code[[2]]$Code, Country_code[[2]]$`Nation/Team`)
code_df3 <- data.frame(Country_code[[3]]$Code, Country_code[[3]]$`Nation (NOC)`)

country_code <- rbind(as.matrix(code_df1),as.matrix(code_df2),as.matrix(code_df3))
country_code <- data.frame(country_code)
colnames(country_code) <- c("CODE", "NATION")


# Extracting country population table from webpage.  


url3 <- "https://en.wikipedia.org/wiki/List_of_countries_by_population_(United_Nations)"

population.list <- url3 %>% read_html() %>% html_nodes("table")  %>% .[1:1] %>%
  html_table(fill = TRUE) 

population.df <- population.list[[1]]

names(population.df) <- make.names(names(population.df))

# Join first 2 URL data on country code
df2_1 <- left_join(olympic_medal,country_code,by=c("COUNTRY" = "CODE"))



# Join df2_1 and 3rd URL data on country name and create a new data frame df_2_1_3 that has columns for country name, the UN continental region, the population (2016), and the number of gold, silver, bronze, and total medals from the Rio Olympics.
  
population.df$Country.or.area <- str_trim(population.df$Country.or.area) 
df2_1$NATION <- str_trim(df2_1$NATION)

df2_1_3 <- left_join(df2_1, population.df, by=c("NATION" = "Country.or.area")) %>%
           filter(!is.na(Population..1.July.2016..2.)) %>%
           select(NATION, UN.continental.region.1., Population..1.July.2016..2., GOLD, SILVER, BRONZE, TOTAL)

colnames(df2_1_3) <- c("NATION", "REGION", "POP16", "GOLD", "SILVER", "BRONZE", "TOTAL")

head(df2_1_3)

```

b) Grouping by region, please fit a linear regression of total medals on population and create a small dataset with only the variables of region and the R^2 from the regression.

```{r warning=FALSE}

lreg <- group_by(df2_1_3, REGION) %>%
        do (ftreg = lm(TOTAL ~ POP16, data = .))

reg_dataset = glance(lreg, ftreg)[c("REGION","r.squared")]

head(reg_dataset)

```

c) Create a long (or tall) version of the dataset in (a) with a variable indicating the type of medal and a variable showing the number of medals for the country of that type.

```{r warning=FALSE}

tall_dataset <- gather(df2_1_3, GOLD:TOTAL, key="medal_type", value = "count")
head(tall_dataset)

```

